|

### 2025-12-12 manuscript/paper.Rmd
updated.

$$
w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\# \text{ discordant pairs}}{{n \choose 2}}
$$ 
and the number of discordant pairs is equivalent to the number of adjacent swaps needed to get the order of $\hat{r}$ correct.

As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$.  This requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4 with 2 and then 3 with 2). The Kendall's $\tau$ value between these two vectors is $\frac{1}{3}$.  We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 = \frac{2}{3}$ .  This is exactly equivalent to $1 - \frac{\#\text{ adj swaps}}{{n \choose 2}}$.

@@

$$
I_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r})p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
$$

@@

@Shieh1998: Weighted Kendall's Tau

@@

Incorporating $m$ into this function is straightforward for many weighting functions.  However, using weighted Kendall's $\tau$, $w(r, \hat{r}, m)$ is slightly more complex and becomes:

@@

where $H(.)$ is the usual Shannon entropy. Using this framework, values of $I^{\star}_w(r,\hat{r},m)$ near 1 indicated competition structures with high efficacy.  Values near 0.25 indicate tournament structures where all possible rankings are equally likely and values near 0 indicate rankings that are [what is the right word here "Inverse efficacy"?]

>>>>

$$
w(\hat{r},r) = \frac{\rho_\tau + 1}{2} = 1 - \frac{\# \text{ discordant pairs}}{{n \choose 2}}
$$ 
and the number of discordant pairs is equivalent to the number of adjacent swaps needed to get the order of $\hat{r}$ correct.  
As an example, consider $r = (1,2,3,4)$ and $\hat{r} = (1,3,4,2)$.  This requires two adjacent swaps to get $\hat{r}$ to equal $r$ (i.e swap 4 with 2 and then 3 with 2). The Kendall's $\tau$ value between these two vectors is $\frac{1}{3}$.  We would then define the weight between the two vectors to be $(\rho_{\tau} + 1)/2 = \frac{2}{3}$ .  This is exactly equivalent to $1 - \frac{\#\text{ adj swaps}}{{n \choose 2}}$.

@@

$$
I_w(r,\hat{r})= \sum_{r}\sum_{\hat{r}}w(r, \hat{r})p(r,\hat{r})\log \frac {p(r,\hat{r})}{p(r)p(\hat{r})}
$$
# Tiered Ranking

This framework was developed to assess the efficacy of the full ranking of the top $k$ teams generated by different competition formats.  However, it may be of interest to have a competition structure that simply selects the top $k$ teams in any order.  Continuing with the Olympics example, if an organizer only cares about the top 3 teams earning medals *in any order* (i.e. $\hat{r} = \{1,2,3,4\}$ and $\hat{r} = \{3,2,1,4\}$ are equivalent), our framework can easily handle this modifying $\hat{r}$ (and $r$?) so that the top $k$ elements are ordered correctly and tied.  That is, for the Olympic example, if $k$ was 3 and $\hat{r} = \{3,2,1,4\}$, we convert all the elements of $\hat{r} \le k$ to 1 (i.e. $\hat{r}$ becomes $\{1,1,1,4\}$), and then we proceed with the exact same formula.  We refer to this as tiered ranking.

This concept can be extended to multiple tiers, which is applicable to a competition structure such as the premier league where the top 4 teams qualify for the Champions league and the bottom 3 teams are relegated.  While this straight foward to compute, we do not evaluate scenarios such as this in this article.

@@

<!-- @Shieh1998: Weighted Kendall's Tau -->

@@

Incorporating $m$ into this function is straightforward for many weighting functions.  We use the weighted Kendall's $\tau$ presented in @Shieh1998 with a slight modification to the sign function. Specifically, in the special case that $sign(0) = 1$.  Then $w(r, \hat{r}, m)$ becomes:

@@

where $H(.)$ is the usual Shannon entropy. Using this framework, values of $I^{\star}_w(r,\hat{r},m)$ near 1 indicated competition structures with high efficacy.  Values near 0.25 indicate tournament structures where all possible rankings are equally likely and values near 0 indicate rankings that are inefficacious (e.g. complete rank reversal).
